---
title: "Descriptive Statistics"
output:
  pdf_document:
    highlight: tango
    toc: yes
  html_notebook:
    toc: yes
---

Import needed Libraries

```{r, message=F, warning=F}

# library(sqldf)
library(tidyverse)
library(naniar)
#library(psych)
library(plotrix)
library(ggplot2)
library(pastecs)
library(ggpubr)
library(car)
# library(userfriendlyscience)
library(gmodels)
library(DescTools)
library(stringr)
library(janitor)
library(dplyr)
library(scales)

```

Import data from CSV. There are many blank values which correspond to 
missing data.  We will replace the blank values with NA so R knows 
they are missing values


```{r}
student_data = read.csv("C:\\Users\\yaleq\\OneDrive - csulb\\Yale_Quan_Thesis\\R_Code\\student_data_2.csv", 
                        header = TRUE, na.strings=c(""," ","NA"))



student_data = read.csv("D:\\OneDrive - csulb\\Yale_Quan_Thesis\\R_Code\\student_data_2.csv", 
                        header = TRUE, na.strings=c(""," ","NA"))

```

Timely Graduation Coding Check

```{r}
student_data$timely_grad = as.factor(student_data$timely_grad)
student_data$graduated_flag = as.factor(student_data$graduated_flag)

levels(student_data$timely_grad)
levels(student_data$graduated_flag)

```

The purpose of this study is to analyze the four and six year graduation rates.  
Graduation rates are calculated based on semesters enrolled (8 semesters is 4 
year, and grater than 8 and up to 12 semesters is 6 years)  Students with a 
reported HS GPA of 0.00 were also excluded and should be investigated as to why 
these numbers were reported

We also want to filter out students who ever enrolled in Winter or 
Summer Classes.  They can be studied later. 

This will be the working data for analysis

```{r}
# count number of students who exceeded the 
# 12 semesters for 6 year grad by cohort

student_data %>% 
  filter(semesters > 12 & winter_summer_flag == 0) %>%
  group_by(cohort) %>%
  summarize(n_students_semester_over_12 = n())
```

```{r}
# count number of students who enrolled 
# in Winter/Summer by cohort

student_data %>% 
  filter(winter_summer_flag == 1) %>%
  group_by(cohort) %>%
  summarize(n_students_winter_or_summer = n())
```

Drop students who are excluded based in time enrolled, winter/summer enrollment

```{r}
full_cohort = student_data %>% filter( semesters < 13 & 
                                         High_School_GPA > 0 & 
                                         winter_summer_flag == 0 & 
                                         cohort >= 2009 & 
                                         start_college != "COB"& 
                                         !is.na(start_college))
```

Cleanup. Drop Columns not being used

```{r}

full_cohort = select(full_cohort, -c(2, 4:8, 31:43, 57:76, 80, 96:108, 121:133,
                                     151:163, 176:188, 201:213, 232:244))

```

Create average rolling difference column

```{r}
full_cohort$mean_rolling_difference = rowMeans(select(full_cohort,c(98:108)), 
                                               na.rm  = TRUE)
```

Export the working dataset for other analysis

```{r}
write.csv(full_cohort,
          "C:\\Users\\yaleq\\OneDrive - csulb\\Yale_Quan_Thesis\\R_Code\\full_cohort.csv",
          row.names = TRUE)

write.csv(full_cohort,
          "D:\\OneDrive - csulb\\Yale_Quan_Thesis\\R_Code\\full_cohort.csv",
          row.names = TRUE)
```


Create a data frame of variables for reference and missing percentages

```{r}
variables = colnames(full_cohort)
variables = as.data.frame(variables)

missing = full_cohort %>% summarize_all(funs(sum(is.na(.)) / length(.)*100))

missing <- as.data.frame(t(as.matrix(missing)))

# Un-comment for CSV files of the variable names and missing percentages

#write.csv(variables,"C:\\Users\\yaleq\\OneDrive - csulb\\Yale_Quan_Thesis\\variables_2.csv",
#          row.names = TRUE)
#write.csv(missing,"C:\\Users\\yaleq\\OneDrive - csulb\\Yale_Quan_Thesis\\missing_2.csv", 
#          row.names = TRUE)
#write.csv(full_cohort,"C:\\Users\\yaleq\\OneDrive - csulb\\Yale_Quan_Thesis\\full_cohort.csv",
#          row.names = TRUE)
```


Create a graduated subset

```{r}
four_year = full_cohort %>% filter(timely_grad == "Four Year Grad")
six_year = full_cohort %>% filter(timely_grad == "Six Year Grad")
did_not_grad = full_cohort %>% filter(timely_grad == "Did not Graduate")
```

Create cohort dataframes

```{r}
Cohort_2009 = full_cohort %>% filter(cohort == '2009')
Cohort_2010 = full_cohort %>% filter(cohort == '2010')
Cohort_2011 = full_cohort %>% filter(cohort == '2011')
Cohort_2012 = full_cohort %>% filter(cohort == '2012')


```

Export to CSV for MATLAB and SAS

```{r}
write.csv(Cohort_2009,"C:\\Users\\yaleq\\Onedrive\\Desktop\\Cohort_2009.csv", 
          row.names = FALSE)
write.csv(Cohort_2010,"C:\\Users\\yaleq\\Onedrive\\Desktop\\Cohort_2010.csv", 
          row.names = FALSE)
write.csv(Cohort_2011,"C:\\Users\\yaleq\\Onedrive\\Desktop\\Cohort_2011.csv", 
          row.names = FALSE)
write.csv(Cohort_2012,"C:\\Users\\yaleq\\Onedrive\\Desktop\\Cohort_2012.csv", 
          row.names = FALSE)
```


Count the number of students in each cohort

```{r}
nrow(Cohort_2009) # Size of 2009 cohort
nrow(Cohort_2010) # Size of 2010 cohort
nrow(Cohort_2011) # Size of 2011 cohort
nrow(Cohort_2012) # Size of 2012 cohort
```

###############################################################################

Graduation Rates

###############################################################################

```{r}
Cohort_2009 %>% 
  group_by(timely_grad) %>%
  summarize(n_2009_students = n())%>%
  mutate(freq = n_2009_students / sum(n_2009_students) * 100)

Cohort_2010 %>% 
   group_by(timely_grad) %>%
   summarize(n_2010_students = n())%>%
   mutate(freq = n_2010_students / sum(n_2010_students) * 100)

Cohort_2011 %>% 
   group_by(timely_grad) %>%
   summarize(n_2011_students = n())%>%
   mutate(freq = n_2011_students / sum(n_2011_students) * 100)

Cohort_2012 %>% 
   group_by(timely_grad) %>%
   summarize(n_2012_students = n())%>%
   mutate(freq = n_2012_students / sum(n_2012_students) * 100)

```

###############################################################################

Major Changes

###############################################################################

Calculate the number and percentage of students in each major change category 

```{r}
full_cohort %>% 
                group_by(cohort, major_change_levels) %>%
                summarize(n_students = n())%>%
                mutate(freq = n_students / sum(n_students) * 100)
```

Calculate the mean, median, and standard deviation of the 
Number of Major Changes for each cohort

```{r}
full_cohort %>% 
   group_by(cohort) %>%
   summarize(min(Number_of_Major_Changes), 
             max(Number_of_Major_Changes), 
             mean(Number_of_Major_Changes), 
             sd(Number_of_Major_Changes))
```

###############################################################################

High School GPA

###############################################################################

```{r}

full_cohort %>% 
   group_by(cohort) %>%
   summarize(min(High_School_GPA), 
             max(High_School_GPA), 
             mean(High_School_GPA), 
             sd(High_School_GPA))

```

###############################################################################

Eligibility Index Scores

###############################################################################

```{r}
full_cohort %>% 
   group_by(cohort) %>%
   summarize(min(Eligibility_Index), 
             max(Eligibility_Index), 
             mean(Eligibility_Index), 
             sd(Eligibility_Index))
```

###############################################################################

CSULB GPA

###############################################################################

Calculate the mean, minimum, maxim, and standard deviation of the students 
overall GPA for each cohort

```{r}
full_cohort %>% 
   group_by(cohort) %>%
   summarize(min(Overall_GPA), max(Overall_GPA), mean(Overall_GPA), sd(Overall_GPA))
```

###############################################################################

Gender

##############################################################################

Count the number of Male and Female students in each cohort (0 = Male)

```{r}
Cohort_2009 %>% 
  group_by(gender_code) %>%
  summarize(n_2009 = n()) %>%
   mutate(freq = n_2009 / sum(n_2009))

Cohort_2010 %>% 
  group_by(gender_code) %>%
  summarize(n_2010 = n()) %>%
  mutate(freq = n_2010 / sum(n_2010))

Cohort_2011 %>% 
  group_by(gender_code) %>%
  summarize(n_2011 = n())%>%
  mutate(freq = n_2011 / sum(n_2011))

Cohort_2012 %>% 
  group_by(gender_code) %>%
  summarize(n_2012 = n())%>%
  mutate(freq = n_2012 / sum(n_2012))

full_cohort %>% 
  group_by(gender_code) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n))
```

###############################################################################

Undeclared Semesters

##############################################################################

Find the average, max, and standard deviation of the number of semesters a 
student is undeclared before declaring their first major

```{r}
full_cohort %>%
  group_by(cohort) %>%
  summarise(avg_sem = ceiling(mean(Semesters_Undeclared)), max_sem = max(Semesters_Undeclared), std_sem = sd(Semesters_Undeclared))

full_cohort %>%
  summarise(avg_sem = ceiling(mean(Semesters_Undeclared)), max_sem = max(Semesters_Undeclared), std_sem = sd(Semesters_Undeclared))


full_cohort %>% 
  group_by(cohort, Semesters_Undeclared) %>%
  summarize(n = n())%>%
  mutate(freq = round(n/ sum(n) * 100,2))
```

###############################################################################

STEM Major at admission

##############################################################################

Count the number of STEM majors at admission and the percentage of the cohort

1 = STEM Major

```{r}

full_cohort$STEM_Entrant = as.factor(full_cohort$STEM_Entrant)

full_cohort %>%
  group_by(cohort, STEM_Entrant) %>%
  summarize(n_students = n())%>%
  mutate(freq = n_students / sum(n_students))


full_cohort %>%
  group_by(STEM_Entrant) %>%
  summarize(n_students = n())%>%
  mutate(freq = n_students / sum(n_students))
```

##############################################################################

Parent Education Level

##############################################################################


Count the number of students at each parent education level for each cohort

```{r}
full_cohort %>%
  group_by(cohort, parent_education_1_code) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n))

full_cohort %>%
  group_by(cohort, parent_education_2_code) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n))
```

##############################################################################

Pell Eligible Students

##############################################################################


Count the number of Pell Eligible students in each cohort and create a bar plot.

```{r}
full_cohort %>%
  group_by(cohort, pell_eligibility_coded) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n) * 100)

full_cohort %>%
  group_by(pell_eligibility_coded) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n) * 100)

```

##############################################################################

Local Preference

##############################################################################

Count the number of students who were admitted with local preference

```{r}
full_cohort %>%
  group_by(cohort, local_coding) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n) * 100)

full_cohort %>%
  group_by(local_coding) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n) * 100)

```


##############################################################################

Minority Status

##############################################################################

Count the number of students who were considered a minority at admissions 
and create a bar chart. 0 = not minority 

```{r}

full_cohort$minority_coding = as.factor(full_cohort$minority_coding)


full_cohort %>%
  group_by(cohort, minority_coding) %>%
  summarize(n = n())%>%
  mutate(freq = n/ sum(n) * 100)

```

##############################################################################

Number of DFW courses

##############################################################################

DFW Classes
```{r}
full_cohort %>%
  group_by(cohort) %>%
  summarise(min_DFW = min(Total_DFW), 
            max_DFW = max(Total_DFW), 
            avg_DFW = ceiling(mean(Total_DFW)), 
            std_DFW = ceiling(sd(Total_DFW)))
```

##############################################################################

Rolling Difference

##############################################################################

Rolling Difference
```{r}
full_cohort %>%
  group_by(cohort) %>%
  summarise(min_roll = min(mean_rolling_difference, na.rm = TRUE), 
            max_roll = max(mean_rolling_difference, na.rm = TRUE), 
            avg_roll = mean(mean_rolling_difference, na.rm = TRUE), 
            std_roll = sd(mean_rolling_difference, na.rm = TRUE))

full_cohort %>%
  summarise(min_roll = min(mean_rolling_difference, na.rm = TRUE), 
            max_roll = max(mean_rolling_difference, na.rm = TRUE), 
            avg_roll = mean(mean_rolling_difference, na.rm = TRUE), 
            std_roll = sd(mean_rolling_difference, na.rm = TRUE))


```

##############################################################################

Starting College

##############################################################################

Calculate the percentage of starting college

```{r}
Cohort_2009 %>% 
  group_by(	start_college) %>%
  summarize(n_students = n())%>%
  mutate(freq = n_students / sum(n_students) * 100) %>%
  mutate(cohort = "2009") %>%
  arrange(desc(n_students, .by_group = TRUE))

Cohort_2010 %>% 
   group_by(start_college) %>%
   summarize(n_students = n())%>%
   mutate(freq = n_students / sum(n_students)) %>%
   mutate(cohort = "2010") %>%
  arrange(desc(n_students, .by_group = TRUE))

Cohort_2011 %>% 
   group_by(start_college) %>%
   summarize(n_students = n())%>%
   mutate(freq = n_students / sum(n_students)) %>%
   mutate(cohort = "2011") %>%
  arrange(desc(n_students, .by_group = TRUE))

Cohort_2012 %>% 
   group_by(start_college) %>%
   summarize(n_students = n())%>%
   mutate(freq = n_students / sum(n_students)) %>%
   mutate(cohort = "2012") %>%
  arrange(desc(n_students, .by_group = TRUE))
```














